Continuous developmental change explains discontinuities in word learning

### 2019-08-13 journal/ms.Rmd
new stuff jounral

---
title             : "Continuous developmental change can explain discontinuities in word learning"
shorttitle        : "Continuous Development of Word Learning"
#numbersections: true

@@

 "Cognitive development is often characterized in term of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using simpler testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model describing how development may proceed in a continuous fashion across the lifespan.  The model accounts for previously documented facts and provides new predictions.  We collected data from preschool children and adults, and we showed that the model can explain various patterns of learning both within the same age and across development. The findings suggest that major aspects of cognitive development that are typically thought of as discontinuities may emerge from simpler, continuous mechanisms."

@@

Cognitive development is sometimes characterized in terms of a succession of discontinuous stages [@piaget1954]. Although intuitively appealing, stage theories can be challenging to integrate with theories of learning, which typically posit that knowledge and skills improve incrementally with experience. Indeed, one of the central challenges of cognitive development has been to explain transitions between stages which appear to be qualitatively different [@carey2009].

Nevertheless, at least in some cases, development may only appear to be stage-like. This appearance can be due, for example, to the use of a cognitively-demanding task which may mask learning, or to the use of statistical thresholding (in particular, p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model---involving similar representations across the lifespan---may provide a simpler and more transparent account of development.

@@

Though infants around 14-month old can distinguish perceptually similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects in the switch task. This failure has initially been taken as evidence that 14-month olds do not encode subtle sounds during meaning learning [@stager1997; @pater2004]. This interpretation suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children, around 17 months, typically do [@werker2002].

@@

The initial discontinuous interpretation has been challenged by subsequent work. For instance, @yoshida2009 investigated whether failure in the switch task reflects a lack of sound encoding during *familiarization*, or whether it is only due to the demands of the *testing* method which does not allow learning below a certain threshold to be detected. They used the same familiarization procedure as @stager1997, but instead of comparing the looking times in 'same' and 'switch' trials, they tested infants using a two-alternative choice task comparing fixations to target and distractor objects [@golinkoff1987; @fernald2006]. Using this simpler and finer-grained task, the researcher found evidence for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from adult studies. If the mismatch between sound discrimination and word learning is only a stage in early infancy, then this mismatch should disappear by adulthood. Nonetheless, even adults show patterns of learning that mirror those shown by 14-month-olds when the sound contrasts more challenging [@white2013; @pajak2016].

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that the phonological form may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding the phonological form of words in a probabilistic fashion. According to this view, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy representations.

In a probabilistic account, a word can be represented as a probability distribution over sound instances organized in a similarity space. The probability is highest at the most typical sound instance. It decreases as the instance becomes less typical. The precision of the representation can be characterized by how much it tolerates slightly atypical pronunciations. This tolerance is captured formally by the variance of the probability distribution: larger variance indicates higher tolerance and lower precision, whereas smaller variance indicates lower tolerance and higher precision.

This framework can explain several findings. In Stager and Werker's original experiment, children are supposed to associate one label "bih" with object 1 and a second label "dih" with object 2.  Though infants could learn that the label "bih" is a better match to object 1 than "dih", they could still judge the sound "dih" as a plausible instance of the label "bih", thanks to the relatively large variance of the early encoding, and this confusion leads to "failure" in the recognition task (Figure \@ref(fig:illus), top). Though learning is small and is easily disrupted by the Switch task, it can still be detected when less demanding methods are used [@yoshida2009].

The learning accuracy increases (and is detected even by demanding tests such as the Switch) for more distinct word-forms (e.g., "lif" vs. "neem") where the perceptual distance is large relative to the variance (Figure \@ref(fig:illus), left).  Distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

Finally, development can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic representations (Figure \@ref(fig:illus), right). Such an increase in precision renders minimally different sounds less confusing. Importantly, a more precise representation still has a non-zero variance --- Learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrate in adults studies [@white2013; @pajak2016].

@@

The probabilistic account has been put forward to explain patterns of learning and development at the qualitative level. However, it is crucial to have a precise computational instantiation of this account which would allow us to *quantify* its explanatory power. We could find one previous work that attempted to provide such a computational instantiation  (Hofer and Levy, 2017). However, this previous work was designed with the goal of reproducing the results of a specific study [@pajak2016] which focused on explaining the mismatch between speech perception and word learning in adults rather than on exploring the mechanism of development.

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before (i.e., the effect of the referents' similarity on word learning). Crucially, we explored the extent to which development can be understood as a continuous refinement in similar representations across the lifespan.

>>>>

---
title             : "Continuous developmental change explains discontinuities in word learning"
shorttitle        : "Continuous Development of Word Learning"
#numbersections: true

@@

 "Cognitive development is often characterized in term of discontinuities, but these discontinuities can sometimes be apparent rather than actual and can arise from continuous developmental change. To explore this idea, we use as a case study the finding by Stager and Werker (1997) that children's early ability to distinguish similar sounds does not automatically translate into word learning skills. Early explanations proposed that children may not be able to encode subtle phonetic contrasts when learning novel word meanings, thus suggesting a discontinuous/stage-like pattern of development. However, later work has revealed (e.g., through using more precise testing methods) that children do encode such contrasts, thus favoring a continuous pattern of development. Here we propose a probabilistic model describing how development may proceed in a continuous fashion across the lifespan. It characterizes developmental change as improvement in the precision of perceptual encoding. Our model explained previous findings in the literature and provided a new prediction --- the referents' visual similarity modulates word leanring accuracy. The models' predictions were corroborated by human data we collected from both preschool children and adults. The braoder impact of this work is to show that computational models, such as ours, can help us explore the extent to which episodes of cogntive development that are typically thought of as discontinuities may emerge from simpler, continuous mechanisms."

@@

Cognitive development is sometimes characterized in terms of a succession of discontinuous stages [@piaget1954]. Although intuitively appealing, this sort of stage theories can be challenging to integrate with theories of learning, which typically posit that knowledge and skills improve incrementally with experience. Indeed, one of the central challenges of cognitive development has been to explain transitions between stages which appear to be qualitatively different [@carey2009].

Nevertheless, at least in some cases, development may only appear to be stage-like. Some discontinuities may be related to how we measure a specific skill --- some tasks may not allow us to detect successful learning.  Other discontinuities may emerge due to statistical thresholding (e.g., p-value < 0.05) which can create a spurious dichotomy between success and failure in observing a given behavior. In such cases, positing discontinuous stages is unnecessary. Instead, a continuous model --- involving similar representations across the lifespan --- may provide a simpler and more transparent account of development.

@@

Though infants around 14-month old can distinguish perceptually similar sound pairs such as "dih" and "bih", they appear to fail in mapping this pair to two different objects in the switch task. This failure was initially taken as evidence that 14-month olds do not encode subtle sounds during meaning learning [@stager1997; @pater2004]. This interpretation suggested a discontinuous/stage-like pattern of development whereby younger children fail to encode the contrastive phonetic detail, whereas older children, around 17 months, typically do [@werker2002].

@@

The initial discontinuous interpretation has been challenged by subsequent work. For instance, @yoshida2009 investigated whether failure in the switch task reflects a lack of sound encoding during *familiarization*, or whether it is only due to the demands of the *testing* method which does not allow learning below a certain threshold to be detected. They used the same familiarization procedure as @stager1997, but instead of comparing the looking times in 'same' and 'switch' trials, they tested infants using a two-alternative choice task comparing fixations to target and distractor objects [@golinkoff1987; @fernald2006]. Using this easier and more precise task, the researcher found evidence for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from adult studies. If the mismatch between sound discrimination and word learning is only a stage in early infancy, then this mismatch should disappear by adulthood. Nonetheless, even adults show patterns of learning that mirror those shown by 14-month-olds when the sound contrasts are more challenging [@white2013; @pajak2016].

Some researchers [@swingley2007; @yoshida2009; @pajak2016] proposed that word knowledge may not be encoded in a binary fashion, i.e., it is not the case that children either succeed or fail in encoding minimal contrast when learning the meanings. Rather, they may be encoding this knowledge in a graded fashion (ses also munakata, 2001). According to this view, development does not so much involve a qualitative shift (i.e., a sudden emergence of an ability that did not exist before) as much as it consists in the continuous refinement of initially noisy knowledge.

Many different computational formalisms can represent graded knowledge. Here we use probabilistic models, a formalism that allows both easy examination of internal representations and quantification of the robustness of these representations. Word knowledge can be characterized with a probability distribution over sound instances organized in a similarity space. The probability is highest at the most typical sound instance. It decreases as the instance becomes less typical. The precision of word knowledge can be characterized by whether it tolerates slightly atypical pronunciations. This tolerance is captured formally by the variance of the probability distribution: larger variance indicates higher tolerance and lower precision, whereas smaller variance indicates lower tolerance and higher precision (for an illustration, see Figure \@ref(fig:illus) top and right panels).

This general framework --- in which the precision of word knowledge is characterize with the variance of a probability distribution --- can already provide an intuitive way of thinking about several findings. In particular, unlike the binary view, the probabilistic view allows for the possibility that knowldege can be successful while still being noisy. This new understanding can provide an account for the fact that children show evidence of learning in some testing condition [@yoshida2009], but not in others [@stager1997].

In a word-pair learning paradigm, children are supposed to associate one label, e.g, "bih", with object 1 and a second label, e.g., "dih", with object 2.  Infants may succeed in learning both associations. Nevertheless, the precision with which the words are encoded can still be large, causing their probability distributions to overalp (Figure \@ref(fig:illus), top). The way this knowledge is tested can lead to divering results.

Failure in the Switch task [@stager1997] can be exlained as follows. Children are understood to succeed if they reject the wrong pairing (e.g., "bih" with object 2). However, the overlap means that children could think of "bih" as plausible misprononciation of "dih". They may require a stronger sound mismatch to confidently reject the possibility that the wrong label is just an acceptable deviation from the tyical pronouciation. In contrast, success in the two-alternative choice task [@yoshida2009] occurs precisely becasue children do not have to reject a wrong pairing. Instead, the setting allows them to just choose which objects is a better match to a given label, thus revealing whether they have a prefernce, albeit samll, for the correct pairing.

In addition to explaining behavior in different tasks, the probabilsitic understanding also accounts for different behaviors in the same task. In particular, when the labels are quite distinct in the pereptual space ("lif" vs. "neem"), the distributions do not overlap as much as in the case of similar-sounding words (Figure \@ref(fig:illus), left). This fact means that the learners will have a stronger preference fot the correct label-object pairings and less tolerance for the wrong ones. That said, perceptual distinctiveness can be enhanced even for minimally different sounds when other cues highlight their difference [@rost2009; @rost2010; @thiessen2007; @yeung09; @dautriche2015].

Finally, development can be understood as an increase in the precision (i.e., a decrease in the variance) of the probabilistic representations, leading to a lower overlap between the similar-sounding words (Figure \@ref(fig:illus), right).  Importantly, a more precise representation still has a non-zero variance --- Learning difficulties can still be induced with challenging stimuli or in cognitively demanding situations as was demonstrate in adults studies [@white2013; @pajak2016].

@@

The probabilistic account has been put forward to explain patterns of learning and development at the qualitative level. However, it is crucial to have a precise computational instantiation of this account which would help us 1) test this theoretical hypothesis more directly and 2) identify the particular parameters that are the locus of developmental change.  One previous study attempted to provide such a computational instantiation [@hofer2017]. However, this previous work was designed with the goal of reproducing the results of a specific study [@pajak2016] which focused on explaining the mismatch between speech perception and word learning in adults rather than on exploring the mechanism of development.

The present work proposes a model of word-pair learning based on the probabilistic account. We tested the ability of this model to both *explain* various findings in previous experiments in both children and adults (e.g., the fact that similar words are harder to learn than different words) and to *predict* new learning patterns that have not been tested before (i.e., the effect of the referents' similarity on word learning). The broader impact of this work is to explore the extent to which development can be understood as a continuous refinement in similar representations across the lifespan.

### 2019-08-13 journal/ms.tex
new stuff jounral

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Continuous developmental change can explain discontinuities in word learning},
            pdfauthor={Abdellah Fourtassi, Sophie Regan, \& Michael C. Frank},
            pdfkeywords={word learning, cognitive development, computational modeling},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Continuous developmental change can explain discontinuities in word
learning}
    \author{Abdellah Fourtassi\textsuperscript{1}, Sophie Regan\textsuperscript{1},
\& Michael C. Frank\textsuperscript{1}}
    \date{}

@@

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using simpler testing methods) that children do
encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. The
model accounts for previously documented facts and provides new
predictions. We collected data from preschool children and adults, and
we showed that the model can explain various patterns of learning both
within the same age and across development. The findings suggest that
major aspects of cognitive development that are typically thought of as
discontinuities may emerge from simpler, continuous mechanisms.''

@@

Cognitive development is sometimes characterized in terms of a
succession of discontinuous stages (Piaget, 1954). Although intuitively
appealing, stage theories can be challenging to integrate with theories
of learning, which typically posit that knowledge and skills improve
incrementally with experience. Indeed, one of the central challenges of
cognitive development has been to explain transitions between stages
which appear to be qualitatively different (Carey, 2009).

Nevertheless, at least in some cases, development may only appear to be
stage-like. This appearance can be due, for example, to the use of a
cognitively-demanding task which may mask learning, or to the use of
statistical thresholding (in particular, p-value \textless{} 0.05) which
can create a spurious dichotomy between success and failure in observing
a given behavior. In such cases, positing discontinuous stages is
unnecessary. Instead, a continuous model---involving similar
representations across the lifespan---may provide a simpler and more
transparent account of development.

@@

Though infants around 14-month old can distinguish perceptually similar
sound pairs such as \enquote{dih} and \enquote{bih}, they appear to fail
in mapping this pair to two different objects in the switch task. This
failure has initially been taken as evidence that 14-month olds do not
encode subtle sounds during meaning learning (Pater, Stager, \& Werker,
2004; Stager \& Werker, 1997). This interpretation suggested a
discontinuous/stage-like pattern of development whereby younger children
fail to encode the contrastive phonetic detail, whereas older children,
around 17 months, typically do (Werker, Fennell, Corcoran, \& Stager,
2002).

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the demands of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this simpler and finer-grained task, the researcher found evidence
for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from
adult studies. If the mismatch between sound discrimination and word
learning is only a stage in early infancy, then this mismatch should
disappear by adulthood. Nonetheless, even adults show patterns of
learning that mirror those shown by 14-month-olds when the sound
contrasts more challenging (Pajak, Creel, \& Levy, 2016; White, Yee,
Blumstein, \& Morgan, 2013).

Some researchers (Pajak et al., 2016; Swingley, 2007; Yoshida et al.,
2009) proposed that the phonological form may not be encoded in a binary
fashion, i.e., it is not the case that children either succeed or fail
in encoding minimal contrast when learning the meanings. Rather, they
may be encoding the phonological form of words in a probabilistic
fashion. According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy representations.

In a probabilistic account, a word can be represented as a probability
distribution over sound instances organized in a similarity space. The
probability is highest at the most typical sound instance. It decreases
as the instance becomes less typical. The precision of the
representation can be characterized by how much it tolerates slightly
atypical pronunciations. This tolerance is captured formally by the
variance of the probability distribution: larger variance indicates
higher tolerance and lower precision, whereas smaller variance indicates
lower tolerance and higher precision.

This framework can explain several findings. In Stager and Werker's
original experiment, children are supposed to associate one label
\enquote{bih} with object 1 and a second label \enquote{dih} with object
2. Though infants could learn that the label \enquote{bih} is a better
match to object 1 than \enquote{dih}, they could still judge the sound
\enquote{dih} as a plausible instance of the label \enquote{bih}, thanks
to the relatively large variance of the early encoding, and this
confusion leads to \enquote{failure} in the recognition task (Figure
\ref{fig:illus}, top). Though learning is small and is easily disrupted
by the Switch task, it can still be detected when less demanding methods
are used (Yoshida et al., 2009).

The learning accuracy increases (and is detected even by demanding tests
such as the Switch) for more distinct word-forms (e.g., \enquote{lif}
vs. \enquote{neem}) where the perceptual distance is large relative to
the variance (Figure \ref{fig:illus}, left). Distinctiveness can be
enhanced even for minimally different sounds when other cues highlight
their difference (Dautriche, Swingley, \& Christophe, 2015; Rost \&
McMurray, 2009, 2010; Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic representations
(Figure \ref{fig:illus}, right). Such an increase in precision renders
minimally different sounds less confusing. Importantly, a more precise
representation still has a non-zero variance --- Learning difficulties
can still be induced with challenging stimuli or in cognitively
demanding situations as was demonstrate in adults studies (Pajak et al.,
2016; White et al., 2013).

@@

The probabilistic account has been put forward to explain patterns of
learning and development at the qualitative level. However, it is
crucial to have a precise computational instantiation of this account
which would allow us to \emph{quantify} its explanatory power. We could
find one previous work that attempted to provide such a computational
instantiation (Hofer and Levy, 2017). However, this previous work was
designed with the goal of reproducing the results of a specific study
(Pajak et al., 2016) which focused on explaining the mismatch between
speech perception and word learning in adults rather than on exploring
the mechanism of development.

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). Crucially, we explored the extent to which development
can be understood as a continuous refinement in similar representations
across the lifespan.

>>>>

\documentclass[english,,man,floatsintext]{apa6}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \ifxetex
    \usepackage{mathspec}
  \else
    \usepackage{fontspec}
  \fi
  \defaultfontfeatures{Ligatures=TeX,Scale=MatchLowercase}
\fi
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
% use microtype if available
\IfFileExists{microtype.sty}{%
\usepackage{microtype}
\UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\usepackage{hyperref}
\hypersetup{unicode=true,
            pdftitle={Continuous developmental change explains discontinuities in word learning},
            pdfauthor={Abdellah Fourtassi, Sophie Regan, \& Michael C. Frank},
            pdfkeywords={word learning, cognitive development, computational modeling},
            pdfborder={0 0 0},
            breaklinks=true}
\urlstyle{same}  % don't use monospace font for urls
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[shorthands=off,main=english]{babel}
\else
  \usepackage{polyglossia}
  \setmainlanguage[]{english}
\fi
\usepackage{graphicx,grffile}
\makeatletter
\def\maxwidth{\ifdim\Gin@nat@width>\linewidth\linewidth\else\Gin@nat@width\fi}
\def\maxheight{\ifdim\Gin@nat@height>\textheight\textheight\else\Gin@nat@height\fi}
\makeatother
% Scale images if necessary, so that they will not overflow the page
% margins by default, and it is still possible to overwrite the defaults
% using explicit options in \includegraphics[width, height, ...]{}
\setkeys{Gin}{width=\maxwidth,height=\maxheight,keepaspectratio}
\IfFileExists{parskip.sty}{%
\usepackage{parskip}
}{% else
\setlength{\parindent}{0pt}
\setlength{\parskip}{6pt plus 2pt minus 1pt}
}
\setlength{\emergencystretch}{3em}  % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\setcounter{secnumdepth}{0}
% Redefines (sub)paragraphs to behave more like sections
\ifx\paragraph\undefined\else
\let\oldparagraph\paragraph
\renewcommand{\paragraph}[1]{\oldparagraph{#1}\mbox{}}
\fi
\ifx\subparagraph\undefined\else
\let\oldsubparagraph\subparagraph
\renewcommand{\subparagraph}[1]{\oldsubparagraph{#1}\mbox{}}
\fi

@@

  \title{Continuous developmental change explains discontinuities in word
learning}
    \author{Abdellah Fourtassi\textsuperscript{1}, Sophie Regan\textsuperscript{1},
\& Michael C. Frank\textsuperscript{1}}
    \date{}

@@

\abstract{
``Cognitive development is often characterized in term of
discontinuities, but these discontinuities can sometimes be apparent
rather than actual and can arise from continuous developmental change.
To explore this idea, we use as a case study the finding by Stager and
Werker (1997) that children's early ability to distinguish similar
sounds does not automatically translate into word learning skills. Early
explanations proposed that children may not be able to encode subtle
phonetic contrasts when learning novel word meanings, thus suggesting a
discontinuous/stage-like pattern of development. However, later work has
revealed (e.g., through using more precise testing methods) that
children do encode such contrasts, thus favoring a continuous pattern of
development. Here we propose a probabilistic model describing how
development may proceed in a continuous fashion across the lifespan. It
characterizes developmental change as improvement in the precision of
perceptual encoding. Our model explained previous findings in the
literature and provided a new prediction --- the referents' visual
similarity modulates word leanring accuracy. The models' predictions
were corroborated by human data we collected from both preschool
children and adults. The braoder impact of this work is to show that
computational models, such as ours, can help us explore the extent to
which episodes of cogntive development that are typically thought of as
discontinuities may emerge from simpler, continuous mechanisms.''

@@

Cognitive development is sometimes characterized in terms of a
succession of discontinuous stages (Piaget, 1954). Although intuitively
appealing, this sort of stage theories can be challenging to integrate
with theories of learning, which typically posit that knowledge and
skills improve incrementally with experience. Indeed, one of the central
challenges of cognitive development has been to explain transitions
between stages which appear to be qualitatively different (Carey, 2009).

Nevertheless, at least in some cases, development may only appear to be
stage-like. Some discontinuities may be related to how we measure a
specific skill --- some tasks may not allow us to detect successful
learning. Other discontinuities may emerge due to statistical
thresholding (e.g., p-value \textless{} 0.05) which can create a
spurious dichotomy between success and failure in observing a given
behavior. In such cases, positing discontinuous stages is unnecessary.
Instead, a continuous model --- involving similar representations across
the lifespan --- may provide a simpler and more transparent account of
development.

@@

Though infants around 14-month old can distinguish perceptually similar
sound pairs such as \enquote{dih} and \enquote{bih}, they appear to fail
in mapping this pair to two different objects in the switch task. This
failure was initially taken as evidence that 14-month olds do not encode
subtle sounds during meaning learning (Pater, Stager, \& Werker, 2004;
Stager \& Werker, 1997). This interpretation suggested a
discontinuous/stage-like pattern of development whereby younger children
fail to encode the contrastive phonetic detail, whereas older children,
around 17 months, typically do (Werker, Fennell, Corcoran, \& Stager,
2002).

The initial discontinuous interpretation has been challenged by
subsequent work. For instance, Yoshida, Fennell, Swingley, and Werker
(2009) investigated whether failure in the switch task reflects a lack
of sound encoding during \emph{familiarization}, or whether it is only
due to the demands of the \emph{testing} method which does not allow
learning below a certain threshold to be detected. They used the same
familiarization procedure as Stager and Werker (1997), but instead of
comparing the looking times in \enquote{same} and \enquote{switch}
trials, they tested infants using a two-alternative choice task
comparing fixations to target and distractor objects (Fernald, Perfors,
\& Marchman, 2006; Golinkoff, Hirsh-Pasek, Cauley, \& Gordon, 1987).
Using this easier and more precise task, the researcher found evidence
for learning even in 14-month olds.

Another challenge to the discontinuous account of development came from
adult studies. If the mismatch between sound discrimination and word
learning is only a stage in early infancy, then this mismatch should
disappear by adulthood. Nonetheless, even adults show patterns of
learning that mirror those shown by 14-month-olds when the sound
contrasts are more challenging (Pajak, Creel, \& Levy, 2016; White, Yee,
Blumstein, \& Morgan, 2013).

Some researchers (Pajak et al., 2016; Swingley, 2007; Yoshida et al.,
2009) proposed that word knowledge may not be encoded in a binary
fashion, i.e., it is not the case that children either succeed or fail
in encoding minimal contrast when learning the meanings. Rather, they
may be encoding this knowledge in a graded fashion (ses also munakata,
2001). According to this view, development does not so much involve a
qualitative shift (i.e., a sudden emergence of an ability that did not
exist before) as much as it consists in the continuous refinement of
initially noisy knowledge.

Many different computational formalisms can represent graded knowledge.
Here we use probabilistic models, a formalism that allows both easy
examination of internal representations and quantification of the
robustness of these representations. Word knowledge can be characterized
with a probability distribution over sound instances organized in a
similarity space. The probability is highest at the most typical sound
instance. It decreases as the instance becomes less typical. The
precision of word knowledge can be characterized by whether it tolerates
slightly atypical pronunciations. This tolerance is captured formally by
the variance of the probability distribution: larger variance indicates
higher tolerance and lower precision, whereas smaller variance indicates
lower tolerance and higher precision (for an illustration, see Figure
\ref{fig:illus} top and right panels).

This general framework --- in which the precision of word knowledge is
characterize with the variance of a probability distribution --- can
already provide an intuitive way of thinking about several findings. In
particular, unlike the binary view, the probabilistic view allows for
the possibility that knowldege can be successful while still being
noisy. This new understanding can provide an account for the fact that
children show evidence of learning in some testing condition (Yoshida et
al., 2009), but not in others (Stager \& Werker, 1997).

In a word-pair learning paradigm, children are supposed to associate one
label, e.g, \enquote{bih}, with object 1 and a second label, e.g.,
\enquote{dih}, with object 2. Infants may succeed in learning both
associations. Nevertheless, the precision with which the words are
encoded can still be large, causing their probability distributions to
overalp (Figure \ref{fig:illus}, top). The way this knowledge is tested
can lead to divering results.

Failure in the Switch task (Stager \& Werker, 1997) can be exlained as
follows. Children are understood to succeed if they reject the wrong
pairing (e.g., \enquote{bih} with object 2). However, the overlap means
that children could think of \enquote{bih} as plausible misprononciation
of \enquote{dih}. They may require a stronger sound mismatch to
confidently reject the possibility that the wrong label is just an
acceptable deviation from the tyical pronouciation. In contrast, success
in the two-alternative choice task (Yoshida et al., 2009) occurs
precisely becasue children do not have to reject a wrong pairing.
Instead, the setting allows them to just choose which objects is a
better match to a given label, thus revealing whether they have a
prefernce, albeit samll, for the correct pairing.

In addition to explaining behavior in different tasks, the probabilsitic
understanding also accounts for different behaviors in the same task. In
particular, when the labels are quite distinct in the pereptual space
(\enquote{lif} vs. \enquote{neem}), the distributions do not overlap as
much as in the case of similar-sounding words (Figure \ref{fig:illus},
left). This fact means that the learners will have a stronger preference
fot the correct label-object pairings and less tolerance for the wrong
ones. That said, perceptual distinctiveness can be enhanced even for
minimally different sounds when other cues highlight their difference
(Dautriche, Swingley, \& Christophe, 2015; Rost \& McMurray, 2009, 2010;
Thiessen, 2007; Yeung \& Werker, 2009).

Finally, development can be understood as an increase in the precision
(i.e., a decrease in the variance) of the probabilistic representations,
leading to a lower overlap between the similar-sounding words (Figure
\ref{fig:illus}, right). Importantly, a more precise representation
still has a non-zero variance --- Learning difficulties can still be
induced with challenging stimuli or in cognitively demanding situations
as was demonstrate in adults studies (Pajak et al., 2016; White et al.,
2013).

@@

The probabilistic account has been put forward to explain patterns of
learning and development at the qualitative level. However, it is
crucial to have a precise computational instantiation of this account
which would help us 1) test this theoretical hypothesis more directly
and 2) identify the particular parameters that are the locus of
developmental change. One previous study attempted to provide such a
computational instantiation (Hofer \& Levy, 2017). However, this
previous work was designed with the goal of reproducing the results of a
specific study (Pajak et al., 2016) which focused on explaining the
mismatch between speech perception and word learning in adults rather
than on exploring the mechanism of development.

The present work proposes a model of word-pair learning based on the
probabilistic account. We tested the ability of this model to both
\emph{explain} various findings in previous experiments in both children
and adults (e.g., the fact that similar words are harder to learn than
different words) and to \emph{predict} new learning patterns that have
not been tested before (i.e., the effect of the referents' similarity on
word learning). The broader impact of this work is to explore the extent
to which development can be understood as a continuous refinement in
similar representations across the lifespan.
